---
title: "Hazim"
author: "Hazim"
date: "August 23, 2019"
output: html_document
---
```{r}
##########loading data and essential packages
list.of.packages <- c("expsmooth", "fpp2","fitdistrplus", "logspline","readxl", "xts","forecast", "fma","lmtest", "tseries","Quandl", "fpp","urca", "TSA","tidyverse","gdata","XLConnect")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages,dependencies=TRUE)

library('expsmooth')
library('fpp2')
library('fitdistrplus')
library('logspline')
library('xts')
library('forecast');
library('fma')
library('lmtest')
library('tseries')
library('Quandl')
library('fpp');
library('urca')
library('TSA')
library('readxl')
library('tidyverse')
library('gdata')
library(ggplot2)

set.seed(123123)
dataPath <- "C:\\Users\\akars\\OneDrive\\Desktop\\MScA Spring\\Time Series Analysis\\Project"
data = read.csv("C:\\Users\\akars\\OneDrive\\Desktop\\MScA Spring\\Time Series Analysis\\Project\\Metro_MedianListingPricePerSqft_TopTier.csv",header = TRUE)
```


### Select Cities
```{r}
metros = c('New York, NY',
           'San Francisco, CA',
           'Chicago, IL',
           'Los Angeles-Long Beach-Anaheim, CA',
           'Miami-Fort Lauderdale, FL')

data_1 <- data[data$RegionName %in% metros,]
data_1 <- data_1[,-2]
```

### Transform Dataset & Split into Train & Test set

```{r}
# Transform price data
data_2 <- as.data.frame(t(data_1)[-1,])
colnames(data_2) <- t(data_1)[1,]
data_2 <- apply(data_2, c(1,2), function(x) as.numeric(x))
ts_1 <- ts(data_2, start = c(2010,1), frequency = 12)

ts_1_train = window(ts_1,start =c(2010,1),end = c(2018,7), frequency=12)
ts_1_test = window(ts_1,start =c(2018,8), frequency=12)

```


#### Taking into account the above analysis we do Holt-Winters'
####NY
```{r}
#Holt-Winters' for New York, NY

NY.hw.add <- hw(ts_1_train[,1],seasonal="additive")
NY.hw.mul <- hw(ts_1_train[,1],seasonal="multiplicative") 

summary(NY.hw.add)
summary(NY.hw.mul)

checkresiduals(NY.hw.add)
checkresiduals(NY.hw.mul)
```

##### For New York,Additive model perofrms better

####SF
```{r}
#Holt-Winters' for San Francisco, CA

SF.hw.add <- hw(ts_1_train[,5],seasonal="additive")
SF.hw.mul <- hw(ts_1_train[,5],seasonal="multiplicative") 

summary(SF.hw.add)
summary(SF.hw.mul)

checkresiduals(NY.hw.add)
checkresiduals(NY.hw.mul)

```

##### For San Francisco,Additive model perofrms better

```{r}
#Holt-Winters' for Chicago, IL

CH.hw.add <- hw(ts_1_train[,3],seasonal="additive")
CH.hw.mul <- hw(ts_1_train[,3],seasonal="multiplicative") 

summary(CH.hw.add)
summary(CH.hw.mul)

checkresiduals(CH.hw.add)
checkresiduals(CH.hw.mul)

```

##### For Chicago, additive performs better

####LA
```{r}
#Holt-Winters' for Los Angeles-Long Beacrh-Anaheim, CA

CA.hw.add <- hw(ts_1_train[,2],seasonal="additive")
CA.hw.mul <- hw(ts_1_train[,2],seasonal="multiplicative") 
summary(CA.hw.add)
summary(CA.hw.mul)

checkresiduals(CA.hw.add)
checkresiduals(CA.hw.mul)

```

##### For LA, additive performs better

####MI
```{r}
#Holt-Winters' for Los Angeles-Long Beacrh-Anaheim, MI

MI.hw.add <- hw(ts_1_train[,4],seasonal="additive")
MI.hw.mul <- hw(ts_1_train[,4],seasonal="multiplicative") 

summary(MI.hw.add)
summary(MI.hw.mul)

checkresiduals(MI.hw.add)
checkresiduals(MI.hw.mul)

```

##### For LA, additive performs better

##### The result shows that Holt-Winters' additive model have the lowest RMSE and thus provide a better forecast than others






#### Taking into account the above analysis we do an sarima with regression with lambda =0 for log transformation
####NY
```{r}
sarima.NY<-auto.arima(ts_1_train[,1],lambda = 0)
eacf(ts_1_train[,1])
(EACF1 <- Arima(ts_1_train[,1], order=c(0,1,1),seasonal = c(1,1,1)))
sarres.NY<-sarima.NY$residuals
checkresiduals(sarima.NY)
##the plot still have some unusual behaviour around 2016 and 2018 but behave normally otherwise
#the residual do seems to be fairly normal
##the ACF is within the confidence bound at all lags which suggest the model fits better than the arima

```
####SF
```{r}
sarima.SF<-auto.arima(ts_1_train[,5],lambda = 0)
eacf(ts_1_train[,5])
(EACF1 <- Arima(ts_1_train[,5], order=c(0,2,0),seasonal = c(2,0,1)))
sarres.SF<-sarima.SF$residuals
checkresiduals(sarima.SF)
##the plot behaves normally typical of a good model
#the residual seems to be normally distrbuted
##the ACF is within the confidence bound at all lags which suggest the model fits better than the arima
```
####CH
```{r}
sarima.CH<-auto.arima(ts_1_train[,3],lambda = 0)
eacf(ts_1_train[,3])
(EACF1 <- Arima(ts_1_train[,3], order=c(1,2,1),seasonal = c(0,0,2)))
sarres.CH<-sarima.CH$residuals
checkresiduals(sarima.CH)
##the plot of the residuals shows that residuals are well behaved
#the residual do seems to be  normally distributed
##the ACF is within the confidence bound at all lags which suggest the model fits better than the arima

```
####CA
```{r}
sarima.CA<-auto.arima(ts_1_train[,2],lambda = 0)
eacf(ts_1_train[,2])
(EACF1 <- Arima(ts_1_train[,2], order=c(0,1,1),seasonal = c(1,1,1)))
sarres.CA<-sarima.CA$residuals
checkresiduals(sarima.CA)
##the residuals plot behave normally and the trend is constant overtime.
#the residual do seems to be fairly normal
##the ACF is within the confidence bound at all lags except lag 11 (and this is due to chance) which suggest the model fits better than the arima
##forecast with SARIMA

```
####MI
```{r}
sarima.MI<-auto.arima(ts_1_train[,4],lambda = 0)
eacf(ts_1_train[,4])
(EACF1 <- Arima(ts_1_train[,4], order=c(0,1,1),seasonal = c(1,1,1)))
sarres.MI<-sarima.MI$residuals
checkresiduals(sarima.MI)
##the plot behaves well and exhibits a constant trend
#the residual seem to be normally distributed
##the ACF is within the confidence bound but exceeds at some lag and this is likely due to chance

```



####Comparing Holt-Winters’ and sarima
####NY
```{r}
######comparing Holt-Winters’ and sarima######
stat.NY<-c(NY.hw.mul$aic,sarima.NY$aic,NY.hw.mul$bic,sarima.NY$bic)
mat.NY<-matrix(stat.NY,nrow=2,ncol=2,byrow=F)
row<-c("holt mult","sarima")
column<-c("aic","bic")
colnames(mat.NY)<-column
rownames(mat.NY)<-row
mat.NY

```
####SF
```{r}
######comparing Holt-Winters’ and sarima######
stat.SF<-c(SF.hw.mul$aic,sarima.SF$aic,SF.hw.mul$bic,sarima.SF$bic)
mat.SF<-matrix(stat.SF,nrow=2,ncol=2,byrow=F)
row<-c("holt mult","sarima")
column<-c("aic","bic")
colnames(mat.SF)<-column
rownames(mat.SF)<-row
mat.SF

```
####CH
```{r}
######comparing Holt-Winters’ and sarima######
stat.CH<-c(CH.hw.mul$aic,sarima.CH$aic,CH.hw.mul$bic,sarima.CH$bic)
mat.CH<-matrix(stat.CH,nrow=2,ncol=2,byrow=F)
row<-c("holt mult","sarima")
column<-c("aic","bic")
colnames(mat.CH)<-column
rownames(mat.CH)<-row
mat.CH

```
####CA
```{r}
######comparing Holt-Winters’ and sarima######
stat.CA<-c(CA.hw.mul$aic,sarima.CA$aic,CA.hw.mul$bic,sarima.CA$bic)
mat.CA<-matrix(stat.CA,nrow=2,ncol=2,byrow=F)
row<-c("holt mult","sarima")
column<-c("aic","bic")
colnames(mat.CA)<-column
rownames(mat.CA)<-row
mat.CA

```
####MI
```{r}
######comparing Holt-Winters’ and sarima######
stat.MI<-c(MI.hw.mul$aic,sarima.MI$aic,MI.hw.mul$bic,sarima.MI$bic)
mat.MI<-matrix(stat.MI,nrow=2,ncol=2,byrow=F)
row<-c("holt mult","sarima")
column<-c("aic","bic")
colnames(mat.MI)<-column
rownames(mat.MI)<-row
mat.MI

```
##It can be seen that Holt-Winters’ have the lowest value for both aic and bic, thus sarima is the best model to forecast.


### Extract All the Output for Final Markdown

```{r}
model_hw <- list(NY = NY.hw.add, LA = CA.hw.add, CH = CH.hw.add, MI = MI.hw.add, SF = SF.hw.add) 
model_sarima <- list(NY = sarima.NY, LA = sarima.CA, CH = sarima.CH, MI = sarima.MI, SF = sarima.SF)

forecast_hw <- list(0)
for(i in 1:5){
  forecast_hw[[i]] <- as.data.frame(forecast(model_hw[[i]],h=12))[,1]
}

forecast_sarima <- list(0)
for(i in 1:5){
  forecast_sarima[[i]] <- as.data.frame(forecast(model_sarima[[i]],h=12))[,1]
}

accuracy_hw <- list(0)
for(i in 1:5){
  accuracy_hw[[i]] <- as.data.frame(accuracy(forecast_hw[[i]], ts_1_test[,i]))
}

accuracy_sarima <- list(0)
for(i in 1:5){
  accuracy_sarima[[i]] <- as.data.frame(accuracy(forecast_sarima[[i]], ts_1_test[,i]))
}


saveRDS(model_hw,file=paste(dataPath,"model_hw.rds",sep="\\"))
saveRDS(model_sarima,file=paste(dataPath,"model_sarima.rds",sep="\\"))
saveRDS(forecast_hw,file=paste(dataPath,"forecast_hw.rds",sep="\\"))
saveRDS(forecast_sarima,file=paste(dataPath,"forecast_sarima.rds",sep="\\"))
saveRDS(accuracy_hw,file=paste(dataPath,"accuracy_hw.rds",sep="\\"))
saveRDS(accuracy_sarima,file=paste(dataPath,"accuracy_sarima.rds",sep="\\"))


```